---
title: "3_MR_Summary_Plotter"
author: "Jonathan Chan"
date: "2025-02-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggrepel)
rm(list=ls())

theme_set(theme_classic())
```

This Rmd notebook details the summary plotting of the results from bidirectional , 2-sample MR of HCM-related traits with HCM case/control status.

The 32 GWS SNPs from Tadros et al, 2025 are used as the instruments for the HCM -> phenotype X direction.

## Import

```{r import}

results_folder <- '../3_output/2_MR/'

analyses <- list.dirs(results_folder, recursive=F, full.names=F)
analyses <- analyses[!str_detect(analyses, 'archive|subgroup|DCM')]
#Don't include DCM analyses in overall plot

mainline_results_importer <- function(folder){
  mainline_results <- read.table(str_c(folder, '/',str_match(folder,'\\/([^\\/]+)_to_')[,2], '_mr_results_mainline.tsv')) %>%
    select(-id.exposure, -id.outcome) %>%
  mutate(exposure = ifelse(exposure=='bmi', 'BMI', exposure),
         outcome = ifelse(outcome=='bmi', 'BMI', outcome)) %>%
    mutate(exposure_outcome = str_c(exposure, outcome, sep=' -> '))
  
  return(mainline_results)
}

mr_results <- map(str_c(results_folder, analyses), ~mainline_results_importer(.)) %>% bind_rows()

write_tsv(mr_results, str_c(results_folder, 'summary_MR_results.tsv'))
```
Define the grouping of the phenotypes into different parameter groups here.

```{r}
crf_vars <- c('BMI', 'CRP','DrinksPerWeek','SmokingInitiation','T2D', 'DBP','eGFR')
ecg_vars <- c('PRinterval','PwaveDuration','QRSinterval','QTcInterval','RRinterval','QTinterval')
fibrosis_vars <- c('T1time')
LV_diastolic_vars <- c('longdsr','radialdsr', 'maxLAVi')
RV_vars <- c('minRAAi','maxRAAi', 'RA_Frac_Change','RVEDVI','RVEF','RVESVI','RVSVI')
related_disease_vars <- c('AF', 'HF')
```

```{r}
sensitivity_or_results_importer <- function(folder){
  mainline_results <- read.table(str_c(folder, '/',str_match(folder,'\\/([^\\/]+)_to_')[,2], '_mr_results_sensitivity_oddsratio.tsv')) %>%
    select(-id.exposure, -id.outcome) %>%
  mutate(exposure = ifelse(exposure=='bmi', 'BMI', exposure),
         outcome = ifelse(outcome=='bmi', 'BMI', outcome)) %>%
    mutate(exposure_outcome = str_c(exposure, outcome, sep=' -> '))
  
  return(mainline_results)
}

mr_results_or <- map(str_c(results_folder, analyses), ~sensitivity_or_results_importer(.)) %>% bind_rows() %>%
  filter(method=='Inverse variance weighted') %>%
  mutate(mtc_pval = p.adjust(pval, method='holm')) %>%
  mutate(label = ifelse(mtc_pval <= 0.05, exposure_outcome, '')) %>%
  mutate(group = ifelse(exposure=='HCM', 'HCM -> X', 'X -> HCM')) %>% #Separate a group variable based on whether HCM is the exposure or outcome
    mutate(exposure_phenogroup = case_when( exposure %in% crf_vars ~ 'CRF',
                                        exposure %in% related_disease_vars ~'Related Diseases',
                                        exposure %in% ecg_vars ~ 'ECG',
                                        exposure %in% fibrosis_vars ~ 'Myocardial Fibrosis',
                                        exposure %in% LV_diastolic_vars ~ 'LV Diastolic Function',
                                        exposure %in% RV_vars ~ 'RV Function')) %>%
  mutate(outcome_phenogroup = case_when( outcome %in% crf_vars ~ 'CRF',
                                        outcome %in% related_disease_vars ~'Related Diseases',
                                        outcome %in% ecg_vars ~ 'ECG',
                                        outcome %in% fibrosis_vars ~ 'Myocardial Fibrosis',
                                        outcome %in% LV_diastolic_vars ~ 'LV Diastolic Function',
                                        outcome %in% RV_vars ~ 'RV Function')) %>%
  mutate(nonHCM_phenogroup = case_when(is.na(exposure_phenogroup) ~ outcome_phenogroup, #This creates the phenogroup for the non-HCM variable in the analysis
                                       is.na(outcome_phenogroup) ~ exposure_phenogroup, 
                                       T~NA)) %>%
  mutate(nonHCM_phenogroup=factor(nonHCM_phenogroup, levels=unique(nonHCM_phenogroup)))

write_tsv(mr_results_or, str_c(results_folder, 'summary_MR_results_oddsratio.tsv'))
```

```{r}

perunit_to_perSD_normaliser <- function(input_tb,pooled_sds, type='mr_results'){
    # `pooled_sds` should be a named vector or list where names match `exposure_outcome` values
  # Example: c("BMI -> HCM" = 4.3, "DBP -> HCM" = 11.5)

  if (!type %in% c('mr_results', 'mr_results_or')) {
    stop('Unsupported tibble type')
  }

  # Make sure all the names in pooled_sds are present in the data
  matched_traits <- names(pooled_sds)
  
  # Create a function to fetch the sd based on the exposure_outcome value
  get_sd <- function(trait) {
    if (trait %in% matched_traits) {
      return(pooled_sds[[trait]])
    } else {
      return(1)  # No change if trait is not listed
    }
  }

  # Apply transformation
  if (type == 'mr_results') {
    output_tb <- input_tb %>%
      rowwise() %>%
      mutate(b = b * get_sd(exposure_outcome)) %>%
      ungroup()
  } else if (type == 'mr_results_or') {
    output_tb <- input_tb %>%
      rowwise() %>%
      mutate(b = b * get_sd(exposure_outcome),
             lo_ci = lo_ci * get_sd(exposure_outcome),
             up_ci = up_ci * get_sd(exposure_outcome)) %>%
      ungroup()
  }

  return(output_tb)
}
```

This provides normalisation of BMI and DBP traits to convert their per-unit effect to per-SD effect using the original SD from their publications supplementary information.

```{r}
bmi_dbp_normalisation=T

# Define standard deviations and sample sizes for three studies
sds <- c(11, 12, 11)        # Standard deviations for the three samples
ns <- c(220501, 50649, 458577)      # Corresponding sample sizes

# Calculate the pooled standard deviation using base R
dbp_keaton24_sd <- sqrt(
  sum((ns - 1) * sds^2) / sum(ns - 1)
  )

rm(sds,ns)

#Repeat with BMI
sds <- c(4.32,5.58,3.76,4.72,3.4,4.43,4.19,4.97,4.43,5.37,3.65,4.76)
ns <- c(4652,6954,4385,5891,10190,15522,16833,23150,15738,27262,18525,19285)
bmi_sidorenko24_sd <- sqrt(
  sum((ns - 1) * sds^2) / sum(ns - 1)
  )

rm(sds,ns)

#If variable to normalise is the exposure, divide the beta by SD to convert the effect per-unit exposure to effect per-SD exposure.
#If variable to normalise is the outcome, multiple the beta by SD to convert the effect (units outcome) to effect (SD outcome).
bmi_dbp_sds <- c('DBP -> HCM' = dbp_keaton24_sd, 
                 'BMI -> HCM' = bmi_sidorenko24_sd,
                 'HCM -> DBP' = 1/dbp_keaton24_sd,
                 'HCM -> BMI' = 1/bmi_sidorenko24_sd)

if(isTRUE(bmi_dbp_normalisation)){
  mr_results <- perunit_to_perSD_normaliser(mr_results, bmi_dbp_sds)
  mr_results_or <- perunit_to_perSD_normaliser(mr_results_or, bmi_dbp_sds, type='mr_results_or')
}
```

This provides normalisation of the Thanaj et al, 2022 diastolic measures i.e longitudinal and radial diastolic strain rates into per-SD measures

```{r}
thanaj_dsr_normalisation <- T

longdsr_thanaj22_sd <- 0.6
radialdsr_thanaj22_sd <- 1.9

thanaj_dsr_sds <- c('longdsr -> HCM' = longdsr_thanaj22_sd,
                    'radialdsr -> HCM' = radialdsr_thanaj22_sd,
                    'HCM -> longdsr' = 1/longdsr_thanaj22_sd,
                    'HCM -> radialdsr' = 1/radialdsr_thanaj22_sd)

if(isTRUE(thanaj_dsr_normalisation)){
  mr_results <- perunit_to_perSD_normaliser(mr_results, thanaj_dsr_sds)
  mr_results_or <- perunit_to_perSD_normaliser(mr_results_or, thanaj_dsr_sds, type='mr_results_or')
}
```

## Summary Plotting

This plots a volcano plot of all the MR analyses with Holm's p-value correction for multiple testing correction (controls FWER but with greater power than Bonferroni and also does not assume independence between tests).

```{r}
#Apply multiple testing correction to the tests
mr_results <- mr_results %>%
  mutate(mtc_pval = p.adjust(pval, method='holm')) %>%
  mutate(group = ifelse(exposure=='HCM', 'HCM -> X', 'X -> HCM')) %>% #Separate a group variable based on whether HCM is the exposure or outcome
  mutate(label = ifelse(mtc_pval <= 0.05, exposure_outcome, '')) %>%
  mutate(exposure_phenogroup = case_when( exposure %in% crf_vars ~ 'CRF',
                                        exposure %in% related_disease_vars ~'Related Diseases',
                                        exposure %in% ecg_vars ~ 'ECG',
                                        exposure %in% fibrosis_vars ~ 'Myocardial Fibrosis',
                                        exposure %in% LV_diastolic_vars ~ 'LV Diastolic Function',
                                        exposure %in% RV_vars ~ 'RV Function')) %>%
  mutate(outcome_phenogroup = case_when( outcome %in% crf_vars ~ 'CRF',
                                        outcome %in% related_disease_vars ~'Related Diseases',
                                        outcome %in% ecg_vars ~ 'ECG',
                                        outcome %in% fibrosis_vars ~ 'Myocardial Fibrosis',
                                        outcome %in% LV_diastolic_vars ~ 'LV Diastolic Function',
                                        outcome %in% RV_vars ~ 'RV Function')) %>%
  mutate(nonHCM_phenogroup = case_when(is.na(exposure_phenogroup) ~ outcome_phenogroup, #This creates the phenogroup for the non-HCM variable in the analysis
                                       is.na(outcome_phenogroup) ~ exposure_phenogroup, 
                                       T~NA)) %>%
  mutate(nonHCM_phenogroup=factor(nonHCM_phenogroup, levels=unique(nonHCM_phenogroup)))
```


```{r}
#Output the summary scatter plot of pvalue with effect size
summary_scatter <- ggplot(mr_results,aes(x=b,y=-log10(mtc_pval)))+
  geom_point(aes(shape=group, col=nonHCM_phenogroup))+
  geom_vline(xintercept=0, linetype='dashed', col='black')+
  geom_hline(yintercept=-log10(0.05), linetype='dashed', col='black')+
  geom_text_repel(aes(label=label, col=nonHCM_phenogroup))+
  xlab('Beta (Change in Outcome per Change in Exposure)')+
  ylab('-log10 (Adjusted p-value)')+
  labs(shape='Direction', col='Non-HCM Phenotype Group')+
  labs(title=str_wrap(str_c('Summary plot of 2-sample Mendelian Randomisation for ', nrow(mr_results), ' analyses'))
       #,caption='Labelled if significant after MTC'
       )

print(summary_scatter)
ggsave('../3_output/3_summary_plots/MR_summary_plot.png',summary_scatter, dpi=600, width=9, height=6)

#Note that the RV traits all provide GWAS summary statistics/betas from the inverse-rank-normalised phenotypes. So their phenotype will be measured in SD.
#Likewise with DBP and BMI which are normalised using the SD from the original publication.
#Likweise with Thanaj DSR
```

Also include the plot with 95% confidence interval.
```{r}
#Output the summary scatter plot of pvalue with effect size
summary_scatter_confint <- ggplot(mr_results_or,aes(x=b,y=-log10(mtc_pval)))+
  geom_point(aes(shape=group, col=nonHCM_phenogroup))+
  geom_errorbar(aes(xmin=lo_ci, xmax=up_ci, col=nonHCM_phenogroup))+
  geom_vline(xintercept=0, linetype='dashed', col='black')+
  geom_hline(yintercept=-log10(0.05), linetype='dashed', col='black')+
  geom_text_repel(aes(label=label, col=nonHCM_phenogroup))+
  xlab('Beta (Change in Outcome per Change in Exposure)')+
  ylab('-log10 (Adjusted p-value)')+
  labs(shape='Direction', col='Non-HCM Phenotype Group')+
  labs(title=str_wrap(str_c('Summary plot of 2-sample Mendelian Randomisation for ', nrow(mr_results), ' analyses'))
       ,caption='Error bars = 95% confidence interval'
       )

print(summary_scatter_confint)
ggsave('../3_output/3_summary_plots/MR_summary_plot_confint.png',summary_scatter_confint, dpi=600, width=9, height=6)

#Note that the RV traits all provide GWAS summary statistics/betas from the inverse-rank-normalised phenotypes. So their phenotype will be measured in SD.
```

I also repeat the above summary plot but with the discrete variables transformed from Log(Odds) to Odds Ratio.

This is defunct because plotting odds-ratio is not good when you show both sides (i.e <0 and >0 are not equally well represented.)
```{r}
# #Convert all discrete variables to per-OR (instead of Log(Odds)) exposure and outcome
# discrete_vars <- c('SmokingInitiation','T2D', related_disease_vars, 'HCM', 'DCM')
# 
# #To convert outcome (per-log(OR) to OR), just exponentiate the beta (increase in log(OR) per unit exposure)
# #To convert exposure (per-log(OR) to OR), divide the 
# mr_results_or <- mr_results_or %>%
#   mutate(plot_y = case_when(outcome %in% discrete_vars ~ or,
#                             T~b),
#          plot_y_upci = case_when(outcome %in% discrete_vars~or_uci95,
#                                  T~up_ci),
#          plot_y_loci = case_when(outcome %in% discrete_vars~or_lci95,
#                                  T~lo_ci))
# 
# #Output the summary scatter plot of pvalue with effect size
# summary_scatter_confint_discreteoutcome_OR <- ggplot(mr_results_or,aes(x=plot_y,y=-log10(mtc_pval)))+
#   geom_point(aes(shape=group, col=nonHCM_phenogroup))+
#   geom_errorbar(aes(xmin=plot_y_loci, xmax=plot_y_upci, col=nonHCM_phenogroup))+
#   geom_vline(xintercept=0, linetype='dashed', col='black')+
#   geom_hline(yintercept=-log10(0.05), linetype='dashed', col='black')+
#   geom_text_repel(aes(label=label, col=nonHCM_phenogroup))+
#   xlab('Beta (Increase in Outcome per Increase in Exposure)')+
#   ylab('-log10 (Adjusted p-value)')+
#   labs(shape='Direction', col='Non-HCM Phenotype Group')+
#   labs(title=str_wrap(str_c('Summary plot of 2-sample Mendelian Randomisation for ', nrow(mr_results), ' analyses'))
#        ,caption='Error bars = 95% confidence interval'
#        )
# 
# print(summary_scatter_confint_discreteoutcome_OR)
# ggsave('../3_output/3_summary_plots/MR_summary_plot_confint_discreteoutcomeOR.png',summary_scatter_confint, dpi=600, width=9, height=6)
```


## Forest Plot for Significant Conclusions

```{r}
sig_mr_results_or <- mr_results_or %>%
  filter(exposure_outcome %in% filter(mr_results,mtc_pval <=0.05)$exposure_outcome)

#Create a faceted plot where I have the top plot = per-unit OR and the bottom plot = per-SD OR
if(isTRUE(bmi_dbp_normalisation) & isTRUE(thanaj_dsr_normalisation)){
  persd_exposure_vars <- c(RV_vars, 'BMI','DBP','longdsr','radialdsr')
} else{
  persd_exposure_vars <- c(RV_vars)
}

sig_mr_results_or <- sig_mr_results_or %>%
  mutate(exposure_plot_group = case_when(exposure%in% persd_exposure_vars ~ 'Per-SD Exposure', 
                                       exposure == 'HCM' ~ 'Per-Log(Odds) Exposure',
                                       T~'Per-Unit Exposure'))

sig_mr_results_or <- sig_mr_results_or %>%
  mutate(outcome_plot_group = ifelse(outcome %in% c('HCM',related_disease_vars,'T2D','SmokingInitiation'), 'OR', 'beta')) %>%
  mutate(exposure_outcome=factor(exposure_outcome))

forest_plot1 <- ggplot(sig_mr_results_or)+
  geom_point(aes(x=b, y= exposure_outcome))+
  geom_errorbar(aes(xmin=lo_ci, xmax=up_ci, y=exposure_outcome))+
  geom_vline(xintercept=0, col='black', linetype='dashed')+
  xlab('Change in Outcome per Change in Exposure')+
  ylab('Exposure -> Outcome')+
  labs(col='Non-HCM Phenotype Group')+
  facet_wrap(vars(exposure_plot_group), nrow=3, drop=TRUE, scales='free')

print(forest_plot1)
ggsave('../3_output/3_summary_plots/sig_MR_forestplot.png',forest_plot1, dpi=600, width=9, height=6)
```

## Subgroup Analyses

I perform subgroup analyses for the significant conclusions seen from the initial MR analysis.

```{r}
results_folder <- '../3_output/2_MR/subgroup_analyses/'
analyses <- list.dirs(results_folder, recursive=F, full.names=F)
analyses <- analyses[!analyses %in% c('main_analysis_notsig')]

subgroup_mr_results_or <- map(str_c(results_folder, analyses), ~sensitivity_or_results_importer(.)) %>% bind_rows() %>%
  filter(method=='Inverse variance weighted') %>%
  mutate(label = ifelse(pval <= 0.05, exposure_outcome, '')) %>%
  mutate(group = ifelse(str_detect(exposure, 'HCM'), 'HCM -> X', 'X -> HCM')) %>% #Separate a group variable based on whether HCM is the exposure or outcome
    mutate(exposure_phenogroup = case_when( exposure %in% crf_vars ~ 'CRF',
                                        exposure %in% related_disease_vars ~'Related Diseases',
                                        exposure %in% ecg_vars ~ 'ECG',
                                        exposure %in% fibrosis_vars ~ 'Myocardial Fibrosis',
                                        exposure %in% LV_diastolic_vars ~ 'LV Diastolic Function',
                                        exposure %in% RV_vars ~ 'RV Function')) %>%
  mutate(outcome_phenogroup = case_when( outcome %in% crf_vars ~ 'CRF',
                                        outcome %in% related_disease_vars ~'Related Diseases',
                                        outcome %in% ecg_vars ~ 'ECG',
                                        outcome %in% fibrosis_vars ~ 'Myocardial Fibrosis',
                                        outcome %in% LV_diastolic_vars ~ 'LV Diastolic Function',
                                        outcome %in% RV_vars ~ 'RV Function')) %>%
  mutate(nonHCM_phenogroup = case_when(is.na(exposure_phenogroup) ~ outcome_phenogroup, #This creates the phenogroup for the non-HCM variable in the analysis
                                       is.na(outcome_phenogroup) ~ exposure_phenogroup, 
                                       T~NA)) %>%
  mutate(nonHCM_phenogroup=factor(nonHCM_phenogroup, levels=unique(nonHCM_phenogroup)))
```

Applying normalisation to some of the continuous values to per-SD.

```{r}
bmi_dbp_subgroup_sds <- c(
  'DBP -> HCMsarcneg' = dbp_keaton24_sd, 
  'DBP -> HCMsarcpos' = dbp_keaton24_sd, 
  'BMI -> HCMsarcneg' = bmi_sidorenko24_sd,
  'BMI -> HCMsarcpos' = bmi_sidorenko24_sd)

thanaj_dsr_subgroup_sds <- c(
                    'HCMsarcneg -> longdsr' = 1/longdsr_thanaj22_sd,
                    'HCMsarcpos -> longdsr' = 1/longdsr_thanaj22_sd,
                    'HCMsarcneg -> radialdsr' = 1/radialdsr_thanaj22_sd,
                    'HCMsarcpos -> radialdsr' = 1/radialdsr_thanaj22_sd)

subgroup_mr_results_or <- perunit_to_perSD_normaliser(subgroup_mr_results_or, bmi_dbp_subgroup_sds, type='mr_results_or')
subgroup_mr_results_or <- perunit_to_perSD_normaliser(subgroup_mr_results_or, thanaj_dsr_subgroup_sds, type='mr_results_or')

if(isTRUE(bmi_dbp_normalisation) & isTRUE(thanaj_dsr_normalisation)){
  persd_vars <- c(RV_vars, 'BMI','DBP','longdsr','radialdsr')
} else{
  persd_vars <- c(RV_vars)
}

```



```{r}
subgroup_mr_results_or <- subgroup_mr_results_or %>%
  mutate(outcome_plot_group = case_when(outcome %in% c('HCM','HCMsarcneg','HCMsarcpos',related_disease_vars,'T2D','SmokingInitiation') ~ 'Discrete Outcome',
                                        outcome %in% persd_vars ~ 'Continuous Outcome')) %>%
  mutate(exposure_plot_group = case_when(exposure %in% persd_vars~ 'Per-SD Exposure',
                                         exposure %in% c('HCM','HCMsarcneg','HCMsarcpos',related_disease_vars,'T2D','SmokingInitiation') ~ 'Per-Log(Odds) Exposure')) %>%
  mutate(exposure_outcome=factor(exposure_outcome)) %>%
  mutate(sarcgroup = ifelse(str_detect(exposure_outcome,'sarcneg'), 'Sarcomere-Negative', 'Sarcomere-Positive'))

forest_plot_subgroup_hcm_exposure<- ggplot(filter(subgroup_mr_results_or, str_detect(exposure, 'HCM')))+
  geom_point(aes(x=b, y= forcats::fct_reorder(exposure_outcome, outcome), col=sarcgroup))+
  geom_errorbar(aes(xmin=lo_ci, xmax=up_ci, y=exposure_outcome,col=sarcgroup))+
  geom_vline(xintercept=0, col='black', linetype='dashed')+
  xlab('Change in Outcome (Log(Odds)) for Discrete & SD for Continuous ) per Change in Exposure')+
  ylab('Exposure -> Outcome')+
  labs(col='Sarcomere Subroup',  caption='95% confidence intervals shown')+
  facet_wrap(vars(outcome_plot_group), nrow=3, drop=TRUE, scales='free')

print(forest_plot_subgroup_hcm_exposure)
ggsave('../3_output/3_summary_plots/sig_MR_forestplot_subgroup_hcmexposure.png',forest_plot_subgroup_hcm_exposure, dpi=600, width=9, height=6)

forest_plot_subgroup_hcm_outcome<- ggplot(filter(subgroup_mr_results_or, str_detect(outcome, 'HCM')))+
  geom_point(aes(x=b, y= forcats::fct_reorder(exposure_outcome, exposure), col=sarcgroup))+
  geom_errorbar(aes(xmin=lo_ci, xmax=up_ci, y=exposure_outcome,col=sarcgroup))+
  geom_vline(xintercept=0, col='black', linetype='dashed')+
  xlab('Change in Outcome (Log(Odds)) per Change in Exposure')+
  ylab('Exposure -> Outcome')+
  labs(col='Sarcomere Subgroup',  caption='95% confidence intervals shown')+
  facet_wrap(vars(exposure_plot_group), nrow=3, drop=TRUE, scales='free')

print(forest_plot_subgroup_hcm_outcome)
ggsave('../3_output/3_summary_plots/sig_MR_forestplot_subgroup_hcmoutcome.png',forest_plot_subgroup_hcm_outcome, dpi=600, width=9, height=6)
```


## Misc.

### Double-Check Separate Instruments

```{r}
library(readxl)

double_check_separate_instruments <- function(associations_filepath, reference_IVlist){
  
  phenotype_name <- str_match(associations_filepath, '\\/([A-Za-z_0-9]+)_associations\\.')[,2]
  
  if(any(str_detect(associations_filepath, 'tsv'))){ #i.e if a file path is passed instead of a list of SNPs
  associations_filepath <- read_tsv(associations_filepath, show_col_types=F) %>%
    filter(`P-VALUE` <= 5 * 10^-8) #Filter for only GWS SNPs in the passed in file
  associations_filepath <- associations_filepath$SNPS
} else if (any(str_detect(associations_filepath, 'csv'))){
  associations_filepath <- read_csv(associations_filepath, show_col_types=F) %>% 
    filter(pValue <= 5 * 10^-8) #Filter for only GWS SNPs in the passed in file
  associations_filepath <- associations_filepath$dbSNP
} else if (any(str_detect(associations_filepath, '.xlsx'))){
  associations_filepath <- read_excel(associations_filepath, sheet='Sheet1')
  associations_filepath <- associations_filepath$rsid
}
  
  overlap <- sum(associations_filepath %in% reference_IVlist)
  if(overlap == 0){
    print(str_c(phenotype_name, ' has no overlap with IV list'))
  } else {
    print(str_c(phenotype_name, ' HAS overlap with IV list'))
    overlapping_snps <- associations_filepath[which(associations_filepath %in% reference_IVlist)]
    print(str_c('Overlapping SNPs = ', str_c(overlapping_snps, collapse = ', ')))
  }
  
}

hcm_gws_snps = str_split_1("rs1048302,rs11687178,rs3845778,rs2540277,rs6747402,rs7612736,rs4894803,rs2191446,rs11748963,rs3176326,rs12210733,rs66520020,rs7824244,rs35006907,rs2645210,rs2177843,rs11196085,rs17617337,rs12270374,rs182427065,rs7487962,rs41306688,rs113907726,rs8006225,rs8033459,rs28768976,rs7210446,rs2644262,rs6566955,rs12460541,rs62222424,rs5760054",',')

association_files <- list.files('../1_data/gwas_associations/', pattern='associations', full.names=T)
walk(association_files, ~double_check_separate_instruments(.,hcm_gws_snps))

```


